---
title: "Take-Home_Ex02"
author: "Liaw Ying Ting, Celin"
date: 27 May 2025
date-modified: "last-modified"

format:
  html:
    code-fold: true
    code-tools: true

execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Overview

For this take-home exercise 2, Mini-Challenge 3 will be chosen and the visualisation will be done on this topic. The background of the mini challenge and the questions will be listed below. For this mini challenge, I will only be displaying and addressing **Questions 2 and 3** below.

# Background

Over the past decade, the community of Oceanus has faced numerous transformations and challenges evolving from its fishing-centric origins. Following major crackdowns on illegal fishing activities, suspects have shifted investments into more regulated sectors such as the ocean tourism industry, resulting in growing tensions. This increased tourism has recently attracted the likes of international pop star Sailor Shift, who announced plans to film a music video on the island.

Clepper Jessen, a former analyst at FishEye and now a seasoned journalist for the Hacklee Herald, has been keenly observing these rising tensions. Recently, he turned his attention towards the temporary closure of Nemo Reef. By listening to radio communications and utilizing his investigative tools, Clepper uncovered a complex web of expedited approvals and secretive logistics. These efforts revealed a story involving high-level Oceanus officials, Sailor Shift’s team, local influential families, and local conservationist group The Green Guardians, pointing towards a story of corruption and manipulation.

# Task and Question

Clepper diligently recorded all intercepted radio communications over the last two weeks. With the help of his intern, they have analyzed their content to identify important events and relationships between key players. The result is a knowledge graph describing the last two weeks on Oceanus. Clepper and his intern have spent a large amount of time generating this knowledge graph, and they would now like some assistance using it to answer the following questions.

1.  **Clepper found that messages frequently came in at around the same time each day.**

    -   Develop a graph-based visual analytics approach to identify any daily temporal patterns in communications.

    -   How do these patterns shift over the two weeks of observations?

    -   Focus on a specific entity and use this information to determine who has influence over them.

2.  **Clepper has noticed that people often communicate with (or about) the same people or vessels, and that grouping them together may help with the investigation.**

    -   Use visual analytics to help Clepper understand and explore the interactions and relationships between vessels and people in the knowledge graph.

    -   Are there groups that are more closely associated? If so, what are the topic areas that are predominant for each group?

        -   For example, these groupings could be related to: Environmentalism (known associates of Green Guardians), Sailor Shift, and fishing/leisure vessels.

3.  **It was noted by Clepper’s intern that some people and vessels are using pseudonyms to communicate.**

    -   Expanding upon your prior visual analytics, determine who is using pseudonyms to communicate, and what these pseudonyms are.

        Some that Clepper has already identified include: “Boss”, and “The Lookout”, but there appear to be many more.

        To complicate the matter, pseudonyms may be used by multiple people or vessels.

    -   Describe how your visualizations make it easier for Clepper to identify common entities in the knowledge graph.

    -   How does your understanding of activities change given your understanding of pseudonyms?

4.  **Clepper suspects that Nadia Conti, who was formerly entangled in an illegal fishing scheme, may have continued illicit activity within Oceanus.**

    -   Through visual analytics, provide evidence that Nadia is, or is not, doing something illegal.

    -   Summarize Nadia’s actions visually. Are Clepper’s suspicions justified?

**Reflection Questions**

-   Given the task to develop visualizations for knowledge graphs, did you find that the challenge pushed you to develop new techniques for visual representation?

-   Did you participate in last year’s challenge? If so, did your experience last year help prepare you for this year’s challenge?

-   What was the most difficult part of working on this year’s data and what could have made it more accessible?

# MC3 Kickstarter

## Getting Started

In the code chunk below, `p_load()` of pacman package is used to load the R packages into R environment.

```{r}
pacman::p_load(tidyverse, jsonlite,
               tidygraph, ggraph,
               SmartEDA, lubridate,
               dplyr, igraph,
               grid)

```

## Import knowledge Graph Data

For this exercise, *mc3.json* file will be used. In the code chunk below, `fromJSON()` of jsonlite package is used to import *mc3.json* file into R and save the output object.

```{r}
MC3 <- fromJSON("data/MC3_graph.json")
MC3_schema <- fromJSON("data/MC3_schema.json")
```

## Inspecting Knowledge graph structure

In the code chunk below, `glimpse()` is used to reveal the structure of the *mc3* knowledge graph.

```{r}
glimpse(MC3)
```

## Extracting the edges and nodes tables

Next, `as_tibble()` of tibble package is used to extract the nodes and links tibble data frames from *mc3* tibble dataframe into two separate tibble dataframes called *mc3_nodes* and *mc3_edges* respectively.

```{r}
mc3_nodes <- as_tibble(MC3$nodes)
mc3_edges <- as_tibble(MC3$edges)
```

## Initial EDA

In the code chunk below, `ExpCatViz()` of SmartEDA package is used to reveal the frequency distribution of all categorical fields in *mc3_nodes* tibble dataframe

```{r}
ExpCatViz(data=mc3_nodes,
          col="lightblue")
```

On the other hand, code chunk below uses `ExpCATViz()` of SmartEDA package to reveal the frequency distribution of all categorical fields in *mc3_edges* tibble dataframe

```{r}
ExpCatViz(data=mc3_edges,
          col="lightblue")
```

## Data Cleaning and Wrangling

The code chunk below performs the following data cleaning tasks:

-   convert values in id field into character data type

-   exclude records with `id` value

-   exclude records with similar id values

-   exclude `thing_collected` field

-   save the cleaned tibble dataframe into a new tibble datatable called `mc_nodes_cleaned`

```{r}
mc3_nodes_cleaned <- mc3_nodes %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(id)) %>%
  distinct(id, .keep_all = TRUE) %>%
  select(-thing_collected)
```

## Cleaning and wrangling edges

Next, the code chunk will be used to:

-   rename source and target fields to form_id and to_id respectively

-   convert values in from_id and to_id fields to character data type

-   exclude values in from_id and to_id which is not found in the id field of mc3_nodes_cleaned

-   exclude records whereby from_id and/or to_id values are missing

-   save the cleaned tibble dataframe and call it mc3_edges_cleaned

```{r}
mc3_edges_cleaned <- mc3_edges %>%
  rename(from_id = source,
         to_id = target) %>%
  mutate(across(c(from_id, to_id),
                as.character)) %>%
  filter(from_id %in% mc3_nodes_cleaned$id,
         to_id %in% mc3_nodes_cleaned$id) %>%
  filter(!is.na(from_id), !is.na(to_id))
```

Next, code chunk below will be used to create mapping of character id in `mc3_nodes_cleaned` to row index.

```{r}
node_index_lookup <- mc3_nodes_cleaned %>%
  mutate(.row_id = row_number()) %>%
  select(id, .row_id)
```

Next, the code chunk below will be used to join and convert `from_id` and `to_id` to integer indices. At the same time, we will drop rows with unmatched nodes.

```{r}
mc3_edges_indexed <- mc3_edges_cleaned %>%
  left_join(node_index_lookup, 
            by = c("from_id" = "id")) %>%
  rename(from = .row_id) %>%
  left_join(node_index_lookup, 
            by = c("to_id" = "id")) %>%
  rename(to = .row_id) %>%
  select(from, to, is_inferred, type) %>%
  filter(!is.na(from) & !is.na(to))  
```

Next, the code chunk below is used to subset nodes to only those referenced by edges

```{r}
used_node_indices <- sort(
  unique(c(mc3_edges_indexed$from, 
           mc3_edges_indexed$to)))

mc3_nodes_final <- mc3_nodes_cleaned %>%
  slice(used_node_indices) %>%
  mutate(new_index = row_number())
```

We will then use th code chunk below to rebuild lookup from old index to new index

```{r}
old_to_new_index <- tibble(
  old_index = used_node_indices,
  new_index = seq_along(
    used_node_indices))
```

Lastly, the code chunk below will be used to update edge indices to match new node table

```{r}
mc3_edges_final <- mc3_edges_indexed %>%
  left_join(old_to_new_index, 
            by = c("from" = "old_index")) %>%
  rename(from_new = new_index) %>%
  left_join(old_to_new_index, 
            by = c("to" = "old_index")) %>%
  rename(to_new = new_index) %>%
  select(from = from_new, to = to_new, 
         is_inferred, type)
```

```{r}
glimpse(mc3_edges_final)
```

# Building the tidygraph object

```{r}
mc3_graph <- tbl_graph(
  nodes = mc3_nodes_final,
  edges = mc3_edges_final,
  directed = TRUE
)
```

```{r}
str(mc3_graph)
```

# Visualising the knowledge graph

Several of the ggraph layouts involve randomisation. In order to ensure reproducibility, it is necessary to set the seed value before plotting by using the code chunk below.

```{r}
set.seed(1234)
```

In the code chunk below, ggraph functions are used to create the whole graph

```{r}
ggraph(mc3_graph, 
       layout = "fr") +
  geom_edge_link(alpha = 0.3, 
                 colour = "gray") +
  geom_node_point(aes(color = `type`), 
                  size = 4) +
  geom_node_text(aes(label = type), 
                 repel = TRUE, 
                 size = 1.0) +
  theme_void()
```

# Question 1

For question 1 visualisations, kindly refer to my teammate [Shermainn's](https://isss608-sherm.netlify.app/take-home_ex/take-home_ex02/take-home_ex02) website and I will be building on her answers.

# Question 2

For nodes, the code `filter()` will be used to filter out only *Person and Vessel* from *Entity*, and *Communication* from *Event* and their *relationship.*

After this filter, the nodes only contain:

-   entity nodes whose sub_type is Person or Vessel

-   Event nodes whose sub_type is Communication

-   Any node whose type is Relationship.

`mutate()` adds a new column called idx to the filtered data frame

`select()` extracts 2 columns from *selected_nodes* and give them a new id and rename to *id_map*

For Edges, `filter()` keeps only those where both endpoints are in the *old* column of *id_map.*

`left_join()` matches each row's *to* value against *id_map\$old* and brings in *id_map\$new* for that *to* node and renames it *to_new*. Same thing for the *from* column.

```{r}
selected_nodes <- mc3_nodes_final %>%
  filter(
    (type == "Entity"     & sub_type %in% c("Person", "Vessel")) |
    (type == "Event"      & sub_type == "Communication")      |
    (type == "Relationship")
  ) %>%
  mutate(idx = row_number())

# mapping old_index → new idx
id_map <- selected_nodes %>%
  select(old = new_index, new = idx)

# 2. filter edges so both ends are in selected_nodes
selected_edges <- mc3_edges_final %>%
  filter(from %in% id_map$old, to %in% id_map$old) %>%
  
  # 3. re-index the endpoints
  left_join(id_map, by = c("from" = "old")) %>%
  rename(from_new = new) %>%
  left_join(id_map, by = c("to"   = "old")) %>%
  rename(to_new   = new) %>%
  select(from = from_new, to = to_new, type)
```

Step 1: we proceed to `filter()` only the edges that we want, specifically *sent, received and evidence_for*. The labels for the edges show "sent", "received" and "evidence_for" using `transmute()`.

Step 2: We then build the node table and colour by sub_type which is called *group*. The label for the names/vessels will be placed under the node using *label.*

Step 3: Draw a Viznetwork graph.

```{r mc3_viznetwork_multi_select, message=FALSE, warning=FALSE}
library(visNetwork)

# 1
viz_edges <- selected_edges %>%
  filter(type %in% c("sent", "received", "evidence_for")) %>%
  transmute(
    from   = from,
    to     = to,
    label  = type,      
    arrows = "to"       
  )

# 2 
all_node_ids <- unique(c(viz_edges$from, viz_edges$to))
viz_nodes <- selected_nodes %>%
  filter(idx %in% all_node_ids) %>%
  transmute(
    id    = idx,        
    label = label,      
    group = sub_type   
                        
  )

# 3
visNetwork(viz_nodes, viz_edges, width = "100%", height = "700px") %>%
  visEdges(
    arrows = "to",
    color  = list(color = "#AAAAAA", highlight = "#FF0000"),
    width  = 4            # thicker edge lines
  ) %>%
  visNodes(
    size = 30,            # bigger node circles
    font = list(size = 20) # bigger node labels
  ) %>%
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = FALSE,
    # Here’s the key change: allow multiple groups to be selected at once
    selectedBy       = list(
      variable = "group",
      style    = 'width:200px; height:26px; font-size:14px;'
    )
  ) %>%
  visLegend(
    useGroups = TRUE,
    position  = "right",
    main      = "Node Sub-types"
  ) %>%
  visLayout(randomSeed = 42)
```

```         
```

```{r mc3_viznetwork_multi_select_with_pv_rel, message=FALSE, warning=FALSE}


# 1) Build the “base” edges exactly as before, but include a default color
viz_edges_base <- selected_edges %>%
  filter(type %in% c("sent", "received", "evidence_for")) %>%
  transmute(
    from   = from,
    to     = to,
    label  = type,          # “sent” / “received” / “evidence_for”
    arrows = "to",
    color  = "#AAAAAA"      # default gray for these original edges
  )

# 2) Identify Person/Vessel ↔ Communication links (type = sent or received).
#    For “sent”:   from = Person/Vessel, to = Communication
#    For “received”: from = Communication, to = Person/Vessel
pc_links <- selected_edges %>%
  filter(type %in% c("sent", "received")) %>%
  transmute(
    person = ifelse(type == "sent", from, to),
    comm   = ifelse(type == "sent", to, from)
  )

# 3) Identify Communication → Relationship edges (type = evidence_for)
cr_links <- selected_edges %>%
  filter(type == "evidence_for") %>%
  transmute(
    comm         = from,
    relationship = to
  )

# 4) Join them so that for each (person, comm) we find all (comm → relationship).
#    This creates new edges: Person → Relationship
new_pv_rel_edges <- pc_links %>%
  inner_join(cr_links, by = "comm") %>%
  transmute(
    from   = person,
    to     = relationship,
    label  = "implied",   # or “related”—pick any hover‐label you like
    arrows = "to",
    color  = "blue"       # draw these edges in blue
  ) %>%
  distinct()             # in case multiple duplicates arise

# 5) Combine the “base” edges with our new Person→Relationship edges
viz_edges <- bind_rows(viz_edges_base, new_pv_rel_edges)

# 6) Build the node data frame from all IDs appearing in viz_edges
all_node_ids <- unique(c(viz_edges$from, viz_edges$to))
viz_nodes <- selected_nodes %>%
  filter(idx %in% all_node_ids) %>%
  transmute(
    id    = idx,         # must match viz_edges$from/to
    label = label,       # human‐readable node label
    group = sub_type     # Person, Vessel, Communication, or Relationship
  )

# 7) Plot with visNetwork. Because each edge row has its own 'color',
#    visNetwork will draw original edges in gray and the new Person→Relationship
#    edges in blue. We still supply highlight color for hover, and allow multi‐select.
visNetwork(viz_nodes, viz_edges, width = "100%", height = "700px") %>%
  visEdges(
    color = list(highlight = "#FF0000"),  # when hovering, edge turns red
    width = 4
  ) %>%
  visNodes(
    size = 30,
    font = list(size = 20)
  ) %>%
  visOptions(
    highlightNearest  = TRUE,
    nodesIdSelection  = FALSE,
    selectedBy        = list(
      variable = "group",
      style    = 'width:200px; height:26px; font-size:14px;'
    )
  ) %>%
  visLegend(
    useGroups = TRUE,
    position  = "right",
    main      = "Node Sub-types"
  ) %>%
  visLayout(randomSeed = 42)

```

By selecting the different sub-groups from the drop-down below, we can conclude the following information.

**Colleagues:**

Person and Person:

-   The Accountant and The Middleman

-   The Intern and Lookout

-   Small Fry and Davis

-   Elise and Nadia Conti

-   Liame Thorne and Nadia Conti

-   Boss and Rodriguez

-   Kelly and Sam

-   Boss and Mrs Money

Person and Vessel:

-   Nadia Conti and Neptune

**Friends:**

-   Sam and Kelly

-   The Intern and Lookout

```{r}
# -------------------------------------------------------------------
# Continue from after mc3_edges_final has been created
# -------------------------------------------------------------------

library(dplyr)
library(tidyr)       # for pivot_longer()
library(visNetwork)
library(RColorBrewer)

# 1. Build sender/receiver pairs and their relationship nodes
send_pairs <- mc3_edges_final %>%
  filter(type == "sent") %>%
  rename(sender = from, comm = to) %>%
  select(sender, comm)

recv_pairs <- mc3_edges_final %>%
  filter(type == "received") %>%
  rename(receiver = to, comm = from) %>%
  select(receiver, comm)

rel_edges <- mc3_edges_final %>%
  filter(type %in% c("evidence_for", "is_inferred")) %>%
  rename(rel = to, comm = from) %>%
  select(rel, comm)

# 2. Join to get each (sender, receiver, rel) for every communication event
triplets <- send_pairs %>%
  inner_join(recv_pairs, by = "comm") %>%          
  inner_join(rel_edges, by = "comm") %>%            
  mutate(
    person1 = pmin(sender, receiver),
    person2 = pmax(sender, receiver)
  ) %>%
  distinct(person1, person2, rel, .keep_all = TRUE) %>% 
  select(person1, person2, rel)

# 3. Collect unique node‐IDs used (persons/vessels and relationship nodes)
node_ids_used <- unique(c(triplets$person1, triplets$person2, triplets$rel))

# 4. Build a small lookup of node info for those IDs
nodes_sub <- mc3_nodes_final %>%
  filter(new_index %in% node_ids_used) %>%
  select(new_index, label, type, sub_type)

# 4a. Define groups:
#    - For entity (Person/Vessel) nodes: group = the person's label (unique per entity)
#    - For relationship nodes: group = the relationship subtype (common to many)
nodes_grouped <- nodes_sub %>%
  mutate(
    group = case_when(
      type == "Entity" & sub_type %in% c("Person", "Vessel") ~ label,
      type == "Relationship"                              ~ sub_type
    )
  )

# 4b. Assign one distinct color per entity (Person/Vessel) group
entity_groups <- nodes_grouped %>%
  filter(type == "Entity") %>%
  distinct(group) %>%
  arrange(group)
n_entities <- nrow(entity_groups)
entity_colors <- grDevices::rainbow(n_entities)
entity_map <- tibble(
  group = entity_groups$group,
  color_entity = entity_colors
)

# 4c. Assign one distinct color per relationship subtype group
rel_groups <- nodes_grouped %>%
  filter(type == "Relationship") %>%
  distinct(group) %>%
  arrange(group) 
n_rels <- nrow(rel_groups)
if (n_rels > 0) {
  brewer_base <- brewer.pal(min(n_rels, 8), "Set1")
  rel_colors <- if (n_rels > length(brewer_base)) {
    colorRampPalette(brewer_base)(n_rels)
  } else {
    brewer_base[1:n_rels]
  }
  rel_map <- tibble(
    group = rel_groups$group,
    color_rel = rel_colors
  )
} else {
  rel_map <- tibble(group = character(0), color_rel = character(0))
}

# 4d. Combine color assignments and build nodes_vis
nodes_vis <- nodes_grouped %>%
  left_join(entity_map, by = "group") %>%
  left_join(rel_map,    by = "group") %>%
  transmute(
    id    = new_index,
    label = label,
    group = group,
    color = ifelse(!is.na(color_entity), color_entity, color_rel)
  )

# 5. Build the visNetwork 'edges' data frame
# 5a. Edges between the two people/vessels (the "communication" edge)
edges_comm_vis <- triplets %>%
  transmute(
    from   = person1,
    to     = person2,
    label  = "comm",
    width  = 2,
    arrows = "to"
  )

# 5b. For each triplet, connect both persons to the relationship node (no label text)
edges_rel_vis <- triplets %>%
  pivot_longer(
    cols      = c(person1, person2),
    names_to  = NULL,
    values_to = "person"
  ) %>%
  transmute(
    from   = person,
    to     = rel,
    label  = "",   # blank so "is_inferred"/"evidence_for" does not show
    width  = 1,
    arrows = "to"
  )

# 5c. Combine both sets of edges
edges_vis <- bind_rows(edges_comm_vis, edges_rel_vis)

# 6. Draw the visNetwork with a dropdown filter by 'group'
visNetwork(nodes_vis, edges_vis) %>%
  visNodes(
    borderWidth         = 1,
    borderWidthSelected = 2
  ) %>%
  visEdges(
    arrows = "to",
    color  = list(color = "#AAAAAA", highlight = "#FF0000"),
    font   = list(align = "middle")
  ) %>%
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE,
    selectedBy       = "group"
  ) %>%
  visLegend(
    addNodes = data.frame(
      label = c(entity_map$group, rel_map$group),
      shape = "dot",
      color = c(entity_map$color_entity, rel_map$color_rel),
      title = "Group"
    ),
    useGroups = FALSE
  )


```

```{r warning=FALSE}
library(dplyr)
library(tidytext)
library(wordcloud)

# 1. Extract communication‐event nodes with their new_index and content
comm_nodes <- mc3_nodes_final %>%
  filter(type == "Event", sub_type == "Communication") %>%
  select(event_index = new_index, content)

# 2. Build a mapping from communication events to entities (Organization subtype)
send_map <- mc3_edges_final %>%
  filter(type == "sent") %>%
  transmute(entity_index = from, event_index = to)

recv_map <- mc3_edges_final %>%
  filter(type == "received") %>%
  transmute(event_index = from, entity_index = to)

event_to_entity <- bind_rows(send_map, recv_map)

# 3. Join event_to_entity with comm_nodes, then keep only Organization entities
comm_by_entity <- event_to_entity %>%
  inner_join(comm_nodes, by = "event_index") %>%
  inner_join(
    mc3_nodes_final %>% 
      select(entity_index = new_index, sub_type, label),
    by = "entity_index"
  ) %>%
  filter(sub_type == "Organization") %>%
  select(organization = label, content)

# 4. Collapse all texts per organization
org_texts <- comm_by_entity %>%
  group_by(organization) %>%
  summarize(text = paste(content, collapse = " "), .groups = "drop")

# 5. Tokenize, remove stop‐words, and compute word frequencies per organization
tidy_org_words <- org_texts %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words, by = "word") %>%
  count(organization, word, sort = TRUE)

# 6. Plot one word cloud per organization
org_list <- unique(tidy_org_words$organization)
n_orgs   <- length(org_list)
n_cols   <- 2
n_rows   <- ceiling(n_orgs / n_cols)

par(mfrow = c(n_rows, n_cols), mar = c(0, 0, 2, 0))
for (org in org_list) {
  freq_tbl <- tidy_org_words %>% filter(organization == org)
  if (nrow(freq_tbl) == 0) {
    plot.new()
    title(main = org)
    next
  }
  v <- setNames(freq_tbl$n, freq_tbl$word)
  wordcloud(
    words        = names(v),
    freq         = as.numeric(v),
    max.words    = 100,
    scale        = c(3, 0.5),
    random.order = FALSE
  )
  title(main = org, line = -1)
}
par(mfrow = c(1, 1))

```

# Question 3

```{r mc3_daily_networks_direct_trimmed, message=FALSE, warning=FALSE, fig.width=16, fig.height=28}


daily_comms <- selected_nodes %>%
  filter(type=="Event", sub_type=="Communication") %>%
  mutate(date=as_date(timestamp)) %>%
  select(idx, date)

all_dates <- sort(unique(daily_comms$date)) %>% discard(is.na)

plots <- map(all_dates, function(d) {
  comm_idx <- daily_comms %>% filter(date==d) %>% pull(idx)
  es <- selected_edges %>% filter(type=="sent", to %in% comm_idx) %>% transmute(comm=to, sender=from)
  er <- selected_edges %>% filter(type=="received", from %in% comm_idx) %>% transmute(comm=from, receiver=to)
  dp <- inner_join(es, er, by="comm") %>% transmute(from=sender, to=receiver, type="sent")
  if (nrow(dp)==0) return(NULL)

  nodes_d <- selected_nodes %>%
    filter(idx %in% unique(c(dp$from, dp$to)), sub_type %in% c("Person", "Vessel"))
  dp <- dp %>% filter(from %in% nodes_d$idx, to %in% nodes_d$idx)
  if (nrow(dp)==0) return(NULL)

  nm <- tibble(old=nodes_d$idx, new=seq_len(nrow(nodes_d)))
  edges2 <- dp %>%
    left_join(nm, by=c("from"="old")) %>% mutate(from=new) %>% select(-new) %>%
    left_join(nm, by=c("to"="old")) %>% mutate(to=new) %>% select(-new)
  nodes2 <- nodes_d %>% mutate(idx2=seq_len(nrow(nodes_d)))

  g <- tbl_graph(nodes=nodes2, edges=edges2, directed=TRUE)
  lay <- create_layout(g, layout="fr", niter=500, area=vcount(g)^2, repulse=vcount(g)^1.2)

  ggraph(lay) +
    geom_edge_link(aes(color=type),
                   arrow=arrow(length=unit(4,"mm"), type="closed"),
                   alpha=0.8, width=1) +
    geom_node_point(aes(color=sub_type), size=4) +
    geom_node_text(aes(label=ifelse(sub_type %in% c("Person","Vessel"), label, "")),
                   repel=TRUE, point.padding=unit(0.5,"lines"),
                   box.padding=unit(0.5,"lines"), size=6) +
    scale_edge_color_manual(values=c(sent="#7570B3")) +
    scale_color_manual(values=c(Person="#1b9e77", Vessel="#d95f02")) +
    ggtitle(as.character(d)) +
    theme_void() +
    theme(plot.title=element_text(size=12,hjust=0.5), legend.position="right")
})

wrap_plots(plots, ncol=2) +
  plot_layout(guides="collect") +
  plot_annotation(title="MC3 Communications by Day")

```

## Part A

The pseudonyms used are:

-   Boss

-   The Intern

-   Mrs Money

-   The Lookout

-   The Accountant

-   The Middleman

## Part B

My visualisations help to break down the communication on a daily basis and to draw the connections between the different relationships.

From the figures above, it can be seen that the Boss is mostly always on the receiving end of the communication, with the exception of Mako. The ones who are in direct contact with Boss are Mrs Money, The Middleman, Davis, Liam Thorne, The Accountant and Mako (Vessel).

Mrs Money, however, receives information mainly from The Intern.

The Lookout and The Intern, however, usually communicates in both directions, with The Intern usually on the receiving end of information.

The Lookout also sends informations to Sam.

**There is a clear hierachy when sending information:**

Lookout -\> Intern -\> Mrs Money/The Accountant/The Middleman -\> Boss

Lookout -\> Sam

**Additional visualisations are done below to further investigate the communication between the pseudonyms and real names.**

```{r}
library(dplyr)
library(visNetwork)

# 1. Keep only Person/Vessel nodes, assigning colors by sub_type
nodes <- selected_nodes %>%
  filter(sub_type %in% c("Person", "Vessel")) %>%
  transmute(
    id    = idx,
    label = label,
    color = ifelse(sub_type == "Person", "#1b9e77", "#d95f02")
  )

# 2. Build direct sender→receiver pairs from communication events
es <- selected_edges %>%
  filter(type == "sent") %>%
  transmute(comm = to, sender = from)
er <- selected_edges %>%
  filter(type == "received") %>%
  transmute(comm = from, receiver = to)

# 3. Join on the communication node to get from→to, then count occurrences
dp <- inner_join(es, er, by = "comm") %>%
  transmute(from = sender, to = receiver)

edges <- dp %>%
  count(from, to, name = "count") %>%
  filter(from %in% nodes$id, to %in% nodes$id) %>%
  mutate(width = count, label = as.character(count))

# 4. Draw a visNetwork: node colors by type, edge width and label by count
visNetwork(nodes, edges) %>%
  visEdges(
    arrows = "to",
    color  = list(color = "#AAAAAA", highlight = "#FF0000"),
    font   = list(align = "middle")
  ) %>%
  visOptions(
    highlightNearest  = TRUE,
    nodesIdSelection  = TRUE
  )



```
