---
title: "Hands on Exercise 4 Part 2: Visual Statistical Analysis"
author: "Liaw Ying Ting, Celin"
date: 7 May 2025
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 6.1 Learning Outcome

Gain hands on experience using:

-   ggstatsplot: to create visual graphics with rich statistical information

-   performance package to visualise model diagnostics

-   parameters package to visualise model parameters

# 6.2 Visual Statistical Analysis with ggstatsplot

[**ggstatsplot**]{.underline} is an extension of ggplot 2 package for creating graphics with details from statistical tests

# 6.3 Getting Started

## 6.3.1 Installing and Launching R Packages

```{r}
pacman::p_load(ggstatsplot, tidyverse)
```

## 6.3.2 Importing Data

Importing Exam.csv data using tidyverse packages

```{r}
exam <- read_csv("Exam_data.csv")
```

## 6.3.3 One-sample test: gghistostats() method

In the code chunk below, gghistostats() is used to build an visual of one-sample test on English scores.

::: panel-tabset
### The Plot

```{r}
#| echo: false
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

### The Code

```{r}
#| eval: false
set.seed(1234)

gghistostats(
  data = exam,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```
:::
