---
title: "Hands-on Exercise 1"
author: "Liaw Ying Ting, Celin"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they willl be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting Charts

## Plotting a simple bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar()
```

## Plotting Bar Chart of Science score Distribution

```{r}
ggplot(data=exam_data, aes(x = SCIENCE)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="Yellow", 
                 fill="Blue") +
  ggtitle("Distribution of Science scores")
```

## Plotting a Box Plot of the English scores of the different races

```{r}
ggplot(data=exam_data, 
       aes(y = ENGLISH, 
           x= RACE)) +
  geom_boxplot()                     
```

## Plotting a Scatter Plot of the Science vs Math scores

```{R}
ggplot(data=exam_data, 
       aes(x= SCIENCE, y=MATHS)) +
  geom_point() +
  geom_smooth(size=0.5)
```

## Plotting a Line graph of the scores of the different classes

```{R}
#| echo: false
ggplot(data=exam_data, 
       aes(x = ENGLISH, 
           colour = CLASS)) +
  geom_density()
```

## Plotting a Violin Plot of science scores in the different classes

```{R}
ggplot(data=exam_data, 
       aes(y = SCIENCE, 
           x= CLASS)) +
  geom_violin()
```
