---
title: "Hands-on Exercise 3 Part 2"
date: 30 Apr 2025
date-modified: "last-modified"
author: "Liaw Ying Ting, Celin"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# 4.1 Programming Animated Statistical Graphics with R

-   Creating animated data visualisation by using **gganimate** and **plotly r** packages

-   Learn how to reshape data by using **tidyr** package

-   Process, wrangle and transform data by using **dplyr** package

## 4.1.1 Basic Concepts of animation

-   Plots do not move

-   Many individual plots are built and stitched together as movie frames

-   Each frame is a different plot and subset drives the flow of the animation when stitched back together

## 4.1.2 Terminology

-   Frame: Represents a different point in time or a different category

-   Animation Attributes: Settings that control how the animation behaves

# 4.2 Getting Started

## 4.2.1 Loading the R packages

```{r}
pacman::p_load(readxl, gifski, gapminder,
               plotly, gganimate, tidyverse)
```

## 4.2.2 Importing the data

```{r}
col <- c("Country", "Continent")
globalPop <- read_xls("GlobalPopulation.xls",
                      sheet="Data") %>%
  mutate_at(col, as.factor) %>%
  mutate(Year = as.integer(Year))
```

# 4.3 Animated Data Visualisation: gganimate methods

-   gganimate extends the grammar of graphics as implemented by ggplot2 to include the description of animation

-   transition\_\*() - how data should be spread out and how it relates to itself across time

-   view\_\*() - how the positional scales should change along the animation

-   shadow\_\*() - how data from other points in time should be presented in the given point in time

-   enter\_\*()/exit\_\*() - how new data should appear and how old data should disappear during the course of the animation

-   ease_aes() - how different aesthetics should be eased during transitions

## 4.3.1 Building a static population bubble plot

```{r}
ggplot(globalPop, aes(x = Old, y = Young, 
                      size = Population, 
                      colour = Country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(title = 'Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young') 
```
